Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-06 robdml.tex
move
introplot
tex

@@

\begin{figure}[t]
	\includegraphics[width=1.0\columnwidth]{introdiag.pdf}
	\caption{\oo{Performance plot with aligned conditions.
	(1) robustness and r1 under the same training cost;
	(2) r1 and training cost for reaching similar robustness;
	(3) training cost and robustness for reaching similar r1.}}
\end{figure}

@@

Given an anchor image $X_a$, we sample a positive image $X_p^S$ (in the same
class as the anchor) and a negative image $X_n^S$ (in different class as the
anchor) with a certain sampling strategy (\eg, semi-hard~\cite{facenet}).
%
For convenience we call this triplet as ``source triplet''.
%
Then its \emph{hardness} is defined as
$H(X_a,X_p^S,X_n^S)=d_\phi(X_a,X_p^S)-d_\phi(X_a,X_n^S)$, which is an internal
part of the triplet loss.

@@

>>>>

%\DeclareMathOperator*{\argmax}{arg\,max}
%\DeclareMathOperator*{\argmin}{arg\,min}

@@

\begin{figure}[t]
	\includegraphics[width=1.0\columnwidth]{introplot.pdf}
	\caption{\oo{Performance plot with aligned conditions.
	(1) robustness and r1 under the same training cost;
	(2) r1 and training cost for reaching similar robustness;
	(3) training cost and robustness for reaching similar r1.}}
\end{figure}

@@

Given an anchor image $X_a$, we sample a positive image $X_p^S$ (in the same
class as the anchor) and a negative image $X_n^S$ (in different class as the
anchor) with a certain sampling strategy (\eg, semi-hard~\cite{facenet}).
%
% For convenience we call this triplet as ``source triplet''.
%
Then its \emph{hardness} is defined as
$H(X_a,X_p^S,X_n^S)=d_\phi(X_a,X_p^S)-d_\phi(X_a,X_n^S)$, which is an internal
part of the triplet loss.
%
For convenience we call this as ``source hardness'', denoted as $H_S$.

@@

In this paper, we argue that the hardness of triplet matters in adversarial
training for DML.
%
The model does not collapse with EST or ACT because the expectation of hardness,
\ie, $E[H]$ will be close to zero.
%
While the variance $\text{Var}[H]$ for EST is higher, and that for ACT is lower.

Instead of directly maximizing the source hardness, we propose to stop the
maximization process when the adversarial examples reach a certain level of
hardness in order to avoid model collapse.
%
Based on the same anchor image $X_a$, we sample another positive image $X_p^D$
and negative image $X_n^D$ with a different triplet sampling strategy, and
we call its hardness as ``destination hardness'', denoted as $H_D$.

Then, we want change (increase)
the hardness level from $H_S$ into $H_D$, by finding adversarial examples for
$(X_a, X_p^S, X_n^S)$.
%
We denote the the hardness of adversarial examples as
$\tilde{H}_S=H(X_a{+}r_a, X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as 
follows:
%
\begin{equation}
	%
	\argmin_{r_a,r_p,r_n} \|\min(0, \tilde{H}_S-H_D)\|^2.
	%
\end{equation}
